Towards a Formally Verified EVM in Production Environment

### 2020-04-27 final-3/paper.tex
paper thorough revision

\subsection{EVM in Why3}
The first phase of the framework is to define a formal specification of EVM in Why3 and provide a platform for rigorous verification.  
We develop the EVM specification, following the Ethereum project yellow paper~\cite{wood2014ethereum}. 
More specifically, the EVM implementation is translated into WhyML, 
the programming and specification language of Why3. %Towards the verification, 
Verification conditions can be further generated
based on the pre- and post-condition specification. 
Generated verification goals can be solved directly through the supported 
Satisfiability Modulo Theories (SMT) solvers or go through a sequence of transformations first. 
% Generated verification goals can be split into a set of subgoals or directly be proved through the supported 
% Satisfiability Modulo Theories (SMT) solvers. 
In cases when the automatic SMT solvers cannot deal with, users can resort to interactive theorem provers for the remaining unsolved proof goals.

@@

To this end, the components of the base infrastructure can be easily specified. 
Stack is defined as a list of elements whose type is \texttt{uint256}, aliased by \texttt{machine word}. 
Memory is defined as a function that maps \texttt{machine word} to an option type \texttt{option memory\_content}. 
Since the memory content could be accessed through 256-bit store and 8-bit store instructions, 
\texttt{memory\_content} is defined as an enumeration type including \texttt{Item8} and \texttt{Item256}. 
Similarly, storage is defined as a function that maps \texttt{machine word} to \texttt{machine word}. 
To reflect the implicit change of the machine state, we defined more miscellaneous types. 
For example, we use \texttt{vmstatus}, \texttt{error} and \texttt{return\_type} to capture the virtual machine status, the operation error, 
and the view of the returned result. 
Furthermore, the record type \texttt{machine\_state} is defined to represent the overall machine state 
which consists of stack, memory, storage, program counter, vmstatus, the instruction list, etc.

@@

\subsubsection{Interpreter Definition}\label{sec:interpreter}
The specification of \texttt{interpreter} formalizes the state transition result of different instructions. 
For a specific instruction, the interpreter determines the resultant machine status developing from the current status. 
Some auxiliary functions are defined to make the definition of the interpreter more concise and compact. 
\begin{verbatim}
let interpreter (m: machine_state): machine_state =
let inst = get_inst m in
match inst with 
  | Some (Arith ADD) -> let (st', a) = (pop_stack (m.mac_stack)) in 
    let (st'', b) = ( pop_stack st') in ...
    {m with mac_stack = push_stack st'' (a' + b'); ...}
\end{verbatim}

For example, in the above code snippet of \texttt{Arith ADD} instrcution, 
\texttt{get\_inst} is used to obtain the next instruction to be executed. 
It is obtained from the instruction list following the program counter. 
% Pop and push operations are the most commonly-used manipulations for the state transition of stack. 
Functions \texttt{push\_stack} and \texttt{pop\_stack} are defined to control the pop and push manipulations
for the state transition of stack. 
% For capturing the status transition result of \texttt{Swap} instructions, 
% recursive functions \texttt{fetch} and \texttt{drop} are defined to support more direct manipulation on lists. 
% Another function \texttt{swap\_stack} is defined further for state transition of swap instructions.
With the support of pre-defined auxiliary functions, 
the definition of the interpreter function is essentially comprised of machine status update with regard to the instructions.

@@

\subsection{Running EVM in Production Environment}
%We customizedly defined a series of drivers to %extract WhyML programs into OCaml programs. 
Fig. \ref{fig2} shows the second phase of the framework: deploy the verified EVM, encoded in Why3, in production environments. The deployment is essentially based on a co-compilation framework between OCaml and Rust.

@@

\paragraph{Open Auction contract verification.}
The open auction contract is mainly comprised of three functions:
(1)~Everyone can send their bids through the \texttt{bid} function when the bidding period is not finished.
When the bid sent by one bidder exceed the current recorded highest bid, the auction state including the highestBidder and highestBid would be updated. Then the withdrawal amount of the previous
highest bidder should be increased by the previous highest bid.
(2)~When one bid is beat by another higher raised bid, the previous bid should be returned back to the
corresponding bidder. Bidders can call the \texttt{withdraw} function to get the money/Ether back.
(3)~The auction is ended by the \texttt{auctionEnd} function. If current time is over the auctionEndTime,
then the auction \texttt{end\_state} should be set to True. As the bidding ended, the beneficiary would 
receive the final highest Bid.

In the WhyML specification, \texttt{auction\_status} records the current state of the auction including the current 
highest bidder, the highest bid and the auction ended state. \texttt{auction\_constant} records the beneficiary and the
auctionEndTime and \texttt{auction\_ended} records the final bidder, bid and the beneficiary claimed money/Ether amount.
The properties to be verified are to guarantee the correctness of the functionality. For example,
in the \texttt{auctionEnd} definition, the postcondition specifies the constraints of 
auction ended state and beneficiary claimed amount that the returned result should satisfy.
Complete specification of the three functions can be found at \cite{Examples}.
%\url{https://github.com/Xiyue-Selina/coordination20}. 
The generated verification conditions can be discharged through \emph{alt-ergo} and \emph{cvc4} automatically.

@@

Research interest of blockchain technology has exploded since the inception of Bitcoin. As the popularity of the second generation of blockchain, 
Ethereum, grows, a series of security vulnerabilities have also appeared. 
Since EVM and smart contracts deal directly with the transactions of valuable cryptocurrency units among multiple parties, 
the security of smart contracts and EVM implementations is of paramount importance. To address the security challenges, 
researchers resorted to the techniques of formal methods and program analysis.

%\begin{itemize}
%    \item \textbf{Formalization Foundation of EVM} 
\noindent\textbf{Specification and Verification.}~An executable formal semantics of EVM has been 
created in the K framework by Everett et al.~\cite{hildenbrandt2017kevm}. 
Compared with KEVM with the support
of matching logic for verification, we use Hoare logic, which serves as a good framework for
verification condition specification, to avoid the complex definitions
of the operational semantics. 
A framework to analyze and verify both the runtime safety and the functional correctness of 
Solidity smart contracts in F*
% , a functional programming language aimed at program verification,
was presented in \cite{bhargavan2016formal}.
Hirai~\cite{hirai2017defining} proposed an EVM implementation in Lem, a language that can be compiled for 
a few interactive theorem provers. Then, safety properties of smart contracts can be proved in proof assistants like Isabelle/HOL.
While in our work, we use WhyML for specification and programming, which supports both logical theories 
and programming data structures. Moreover, both automated and interactive external theorem provers 
can be relied on to discharge verification conditions. 
% Generally, the central design choice of the framework is to leverage the infrastructure of 
% Why3, the deductive verification platform where a set of external theorem provers, both automated and interactive, 
% can be relied on to discharge verification conditions. At the same time, 
% OCaml programs can be directly extracted from WhyML programs through the extraction mechanism for further tests.
% Besides the above formalizations of EVM, 
% there are also implementations of EVM in other languages such as %Javascript~\cite{vmjs}, 
% Go~\cite{vmgo} and Ruby~\cite{cryptape}.
%    \item \textbf{Verification of Smart Contracts}

\noindent\textbf{Testing and Debugging.}~The \texttt{hevm} project~\cite{hevm} is implemented in Haskell for unit testing and debugging of smart contracts.
% though the EVM implementation is not completed yet. 
Sergey et al.~\cite{sergey2017concurrent} provided a new perspective between smart contracts and concurrent objects,
 based on which existing tools and insights for understanding, 
 debugging and verifying concurrent objects can be used on smart contract behaviors. 
 In \cite{luu2016making}, several new security problems were pointed out and a way to enhance the operational semantics of 
 Ethereum was proposed to make smart contracts less vulnerable. 
 Due to the difficulty of correcting the semantics of Ethereum, Luu et al.
 ~\cite{luu2016making} also implemented a symbolic execution tool \texttt{OYENTE} to find security bugs.
While in our work, executable OCaml programs can be directly extracted from WhyML programs for further tests 
with the support of extraction mechanism.
%  of the deductive verification platform.

@@

\section{Conclusion}\label{Sec: Conclusion}
We propose a framework to enable formal specification, verification and testing towards EVM. In this framework, the formalization of EVM is specified in WhyML, the programming and specification language of Why3. Based on the EVM formalization, automatic SMT solvers and interactive theorem provers provided in Why3 can be employed for verification. The OCaml implementation of EVM is extracted from the WhyML specification and then glued with Rust implementation based on the coupling framework. The coupling framework provides the interaction mechanism between OCaml and Rust, which allows us to perform tests on the new implementation without additional interface implementation.

>>>>

\subsection{EVM in Why3}
The first phase of the framework is to define a formal specification of EVM in Why3 
and provide a platform for rigorous verification.  
We develop the EVM specification, following the Ethereum project yellow paper~\cite{wood2014ethereum}. 
More specifically, the EVM implementation is translated into WhyML, 
the programming and specification language of Why3. %Towards the verification, 
Verification conditions can be further generated
based on the pre- and post-condition specification. 
Generated verification goals are solved directly through the supported 
solvers or go through a sequence of transformations first. 
% Generated verification goals can be split into a set of subgoals or directly be proved through the supported 
% Satisfiability Modulo Theories (SMT) solvers. 
In cases when the automatic SMT solvers cannot deal with, 
users can resort to interactive theorem provers for the remaining unsolved proof goals.

@@

To this end, the components of the base infrastructure can be specified. 
Stack is defined as a list of elements whose type is \texttt{uint256}, aliased by \texttt{machine word}. 
Memory is defined as a function that maps \texttt{machine word} to an option type \texttt{option memory\_content}. 
% Since the memory content could be accessed through 256-bit store and 8-bit store instructions, 
% \texttt{memory\_content} is defined as an enumeration type including \texttt{Item8} and \texttt{Item256}. 
Similarly, storage is defined as a function that maps \texttt{machine word} to \texttt{machine word}. 
To reflect the implicit change of the machine state, we defined more miscellaneous types. 
For example, we use \texttt{vmstatus}, \texttt{error} and \texttt{return\_type} to 
capture the virtual machine status, the operation error, 
and the view of the returned result. 
Furthermore, the record type \texttt{machine\_state} is defined to represent the overall machine state 
which consists of stack, memory, storage, program counter, vmstatus, the instruction list, etc.

@@

\subsubsection{Interpreter Definition}\label{sec:interpreter}
The specification of \texttt{interpreter} formalizes the state transition result of different instructions. 
For a specific instruction, the interpreter determines the result machine state developing from the current state. 
Some auxiliary functions are defined to make the definition of the interpreter more concise and compact. 
\begin{verbatim}
let interpreter (m: machine_state): machine_state =
let inst = get_inst m in
match inst with 
  | Some (Arith ADD) -> let (st', a) = (pop_stack (m.mac_stack)) in 
    let (st'', b) = (pop_stack st') in ...
    {m with mac_stack = push_stack st'' (a' + b'); ...}
\end{verbatim}

In the above code snippet, \texttt{get\_inst} is used to obtain the next instruction to be executed. 
It is obtained from the instruction list following the program counter. 
In the case of \texttt{Arith ADD} instrcution, the numbers to perform the add operation on are popped out of the stack first and the
result is pushed into the stack after the calculation. As a result, the stack state is updated as a component of the machine state.
% Pop and push operations are the most commonly-used manipulations for the state transition of stack. 
In this process, functions \texttt{push\_stack} and \texttt{pop\_stack} are defined to control the push and pop manipulations
for the state transition of stack. 
% For capturing the status transition result of \texttt{Swap} instructions, 
% recursive functions \texttt{fetch} and \texttt{drop} are defined to support more direct manipulation on lists. 
% Another function \texttt{swap\_stack} is defined further for state transition of swap instructions.
With the support of pre-defined auxiliary functions, 
the definition of the interpreter function is essentially comprised of machine state update with regard to the instructions.

@@

\subsection{Running EVM in Production Environment}
%We customizedly defined a series of drivers to %extract WhyML programs into OCaml programs. 
Fig. \ref{fig2} shows the second phase of the framework: deploy the extracted OCaml implementation from Why3 in production environments. 
The deployment is essentially based on a co-compilation framework between OCaml and Rust.

@@

\paragraph{Open Auction contract verification.}
The open auction contract is mainly comprised of three functions:
(1)~Everyone can send their bids through the \texttt{bid} function when the bidding period is not finished.
When the bid sent by one bidder exceeds the current recorded highest bid, the auction state including the highestBidder and highestBid would be updated. Then the withdrawal amount of the previous
highest bidder should be increased by the previous highest bid.
(2)~When one bid is beaten by another higher raised bid, the previous bid should be returned back to the
corresponding bidder. Bidders can call the \texttt{withdraw} function to get the money/Ether back.
(3)~The auction is ended by the \texttt{auctionEnd} function. If current time is already greater than the auctionEndTime,
then the auction \texttt{end\_state} should be set to True. As the bidding ended, the beneficiary would 
receive the final highest Bid.

In the WhyML specification, \texttt{auction\_status} records the current state of the auction including the current 
highest bidder, the highest bid and the auction ended state. \texttt{auction\_constant} records the beneficiary and the
auctionEndTime and \texttt{auction\_ended} records the final bidder, bid and the beneficiary claimed money/Ether amount.
The properties to be verified are to guarantee the correctness of the functionality. For example,
in the \texttt{auctionEnd} definition, the postcondition specifies the constraints of 
auction ended state and beneficiary claimed amount that the returned result should satisfy.
Complete specification of the functions can be found at \cite{Examples}.
%\url{https://github.com/Xiyue-Selina/coordination20}. 
The generated verification conditions can be discharged through \emph{alt-ergo} and \emph{cvc4} automatically.

@@

Research interest of blockchain technology has exploded since the inception of Bitcoin. 
As the popularity of the second generation of blockchain, 
Ethereum, grows, a series of vulnerabilities have also appeared. 
Since EVM and smart contracts deal directly with the transactions of valuable cryptocurrency units 
among multiple parties, 
the safety of smart contracts and EVM implementations is of paramount importance. 
To address these challenges, 
researchers resorted to the techniques of formal methods and program analysis.

%\begin{itemize}
%    \item \textbf{Formalization Foundation of EVM} 
\noindent\textbf{Specification and Verification.}~An executable formal semantics of EVM has been 
created in the K framework by Everett et al.~\cite{hildenbrandt2017kevm}. 
Compared with KEVM with the support
of matching logic for verification, we use Hoare logic, which serves as a good framework for
verification condition specification, to avoid the complex definitions
of the operational semantics. 
A framework to analyze and verify the safety and the correctness of 
Solidity smart contracts in F*
% , a functional programming language aimed at program verification,
was presented in \cite{bhargavan2016formal}.
Hirai~\cite{hirai2017defining} proposed an EVM implementation in Lem, a language that can be compiled for 
a few interactive theorem provers. Then, safety properties of smart contracts can be proved in proof assistants like Isabelle/HOL.
While in our work, we use WhyML for specification and programming, which supports both logical theories 
and programming data structures. Moreover, both automated and interactive external theorem provers 
can be relied on to discharge verification conditions. 
% Generally, the central design choice of the framework is to leverage the infrastructure of 
% Why3, the deductive verification platform where a set of external theorem provers, both automated and interactive, 
% can be relied on to discharge verification conditions. At the same time, 
% OCaml programs can be directly extracted from WhyML programs through the extraction mechanism for further tests.
% Besides the above formalizations of EVM, 
% there are also implementations of EVM in other languages such as %Javascript~\cite{vmjs}, 
% Go~\cite{vmgo} and Ruby~\cite{cryptape}.
%    \item \textbf{Verification of Smart Contracts}

\noindent\textbf{Testing and Debugging.}~The \texttt{hevm} project~\cite{hevm} is implemented in Haskell for unit testing and debugging of smart contracts.
% though the EVM implementation is not completed yet. 
Sergey et al.~\cite{sergey2017concurrent} provided a new perspective between smart contracts and concurrent objects,
based on which existing tools for understanding and
debugging concurrent objects can be used on smart contract behaviors. 
In \cite{luu2016making}, several new security problems were pointed out and a way to enhance the operational semantics of 
Ethereum was proposed to make smart contracts less vulnerable. 
Due to the difficulty of correcting the semantics of Ethereum, Luu et al.
~\cite{luu2016making} also implemented a symbolic execution tool \texttt{OYENTE} to find security bugs.
While in our work, executable OCaml programs can be directly extracted from WhyML programs for further tests 
with the support of customized drivers and extraction mechanism.
%  of the deductive verification platform.

@@

\section{Conclusion}\label{Sec: Conclusion}
We propose a framework to enable formal specification, verification and testing towards EVM. 
In this framework, the formalization of EVM is specified in WhyML, 
% the programming and specification language of Why3. 
based on which, automatic SMT solvers and interactive theorem provers can be employed for verification. 
The OCaml implementation of EVM is extracted from the WhyML specification and then glued with Rust implementation based on the coupling framework. The coupling framework provides the interaction mechanism between OCaml and Rust, which allows us to perform tests on the new implementation without additional interface implementation.
